Formal Synthesis of Digital Controllers for Observable Continuous Plants with Transient Performance Specifications

### 2017-10-12 TACAS2018/paper.tex
edited the numerical representation sec

In this work, equations~\eqref{eq:ode1} and~\eqref{eq:ode2} is soundly discretized in time (see Section~\ref{sec:cont_aa}) \todo{or use another paper as reference and remove sec .1?} into
\begin{align}
\label{eq:plant}
x_{k+1} = A_d x_k+ B_d u_k\\
y_{k} = C_d x_k + D_d u_k
\end{align} 
%
where $k \in \mathbb N$ and $x_{0}=x(0)$ is the initial state. 
$A_d$, $B_d$, $C_d$ and $D_d$ denote the matrices that describe the discretized plant dynamics, whereas $A$, $B$, $C$ and $D$ denote the continuous plant dynamics.  
%% We synthesize for requirements over this discrete-time domain. 
%% Later, we will address the issue of variable quantization, 
%% as introduced by the ADC/DAC conversion blocks (Fig.~\ref{fig:digitalsystem}).

 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 \subsubsection{Discrete dynamics using reals for systems without inputs}\label{sec:real_discrete_no_inputs}
 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 \begin{theorem}
 Given $\dot{\vec{x}}=\mat{A}\vec{x}$, where $\mat{A}=\mat{S}\mat{J}\mat{S}^{-1}$, the expression
 \begin{align}
 \vec{x}_T&=\vec{x}(t=T)=\mat{A}_{T}\vec{x}_0\\
 \mat{A}_{T}&= \mat{S}
 \left [ \begin{array}{cccc}
 e^{T\lambda_1}  & s_1\frac{T^{1}e^{T\lambda_i}}{(1)!} & \hdots  & s_i\frac{T^{p-1}e^{T\lambda_i}}{(p-i)!} \\
0 & e^{T\lambda_i}  & s_i\frac{T^{j-i}e^{T\lambda_i}}{(j-i)!} & \vdots \\
\vdots & & \ddots & \vdots \\
0 & \cdots & 0  &e^{T\lambda_i} \\
\end{array} \right ]
 \mat{S}^{-1}
 \label{eq:continuous_tube_dyn}\\
 &\text{where } s_i=\left\{\begin{array}{cc}1&gm(\lambda_i)>1\\0&gm(\lambda_i)=1\end{array}\right.,\nonumber
 \end{align}
$\lambda_i \in \mat{J}$ are the eigenvalues of $\mat{A}$, and $gm(\lambda_i)$ is the geometric multiplicity of $\lambda_i$.  $\vec{x}_T$ is a witness of the system $\dot{\vec{x}}(t)=\mat{A}\vec{x}$ at time t=T.
 \end{theorem}
 \begin{corollary}
 If $\mat{J}$ is diagonal and there exists a discrete dynamics matrix for a sampling time $T_d$ such that $A_d=e^{\mat{A} T_d}$, then $\vec{x}_k=A_d^k\vec{x} : k \in (0,\infty)$ is a witness of the system at time $t=kT_d$.
 \end{corollary}
% \begin{proof}
% Let us recall equation \ref{eq:discretize}. The discrete representation of the system dynamics discretised with sample time $T_1$ is ruled by the formula
% $\mat{A}_1 = e^{\mat{A} T_1}$. From matrix theory, we have 
%\begin{align}
% e^{\mat{A}}&=\sum_{k=0}^\infty \frac{1}{k!}\mat{A}^k
%\end{align} 
%\begin{align} 
% e^{\mat{S}\mat{J}\mat{S}^{-1}}&=\sum_{k=0}^\infty \frac{1}{k!}\left(\mat{S}\mat{J}\mat{S}^{-1}\right)^k
% =\mat{S} \left (\sum_{k=0}^\infty \frac{1}{k!}\mat{J}^k\right) \mat{S}^{-1}\nonumber\\
% &=\mat{S}e^{\mat{J}}\mat{S}^{-1}
%\end{align} 
%\begin{equation}
%\mat{A}=\mat{S}\mat{J}\mat{S}^{-1} \Rightarrow \vec{A}_1 = \mat{S}\mat{J}_1\mat{S}^{-1}= \mat{S}e^{\mat{J} T_1}\mat{S}^{-1}.
% \end{equation}
% 
% Let $\mat{J}$ be a diagonal matrix, such that 
% $${\lambda_1}_i=e^{\lambda_i T_1}, \forall \lambda_i \in \mat{J}.$$
% Let us now take a different sample rate $T_2$, such that 
% $${\lambda_2}_i=e^{\lambda_i T_2}, \forall \lambda_i \in \mat{J}.$$
% We can then say that 
% \begin{equation}
% {\lambda_2}_i=e^{\lambda_i T_1 \frac{T_2}{T_1}}={\lambda_1}_i^{\frac{T_2}{T_1}} \Rightarrow A_2=A_1^{\frac{T_2}{T_1}}.
% \end{equation}
% This proves the theorem for $gm(\lambda_i)=1$.
% We now note that given an exponentiated Jordan form with geometric multiplicity, the upper diagonal terms can be
% referenced to the original eigenvalues modified by the sampling rate.
%  Let $\mat{J}\in \mathbb{R}^{p \times p}$ be a Jordan Canonical matrix.
% Then
%\begin{align}
% \mat{J}_1&=\sum_{k=0}^\infty \frac{1}{k!}\left(\mat{J}T_1\right)^k=\sum_{k=0}^\infty \frac{1}{k!} \left [ \begin{array}{cccc}
% \lambda_i^k  & \binom{k}{1}  \lambda^{k-1} & \hdots  & \binom{k}{p-1} \lambda_i^{k-p+1} \\
%& \lambda_i^k  & \binom{k}{1}  \lambda_i^{k-1} & \vdots \\
%\vdots & \ddots & \ddots & \vdots \\
%& &  &\lambda_i^k \\
%\end{array} \right ] T_1^k
%\end{align}
%The eigenvalues remain the same as in the diagonal case, but the upper triangular terms are of the form:
%\begin{align}
%\forall j>i, c_{ij}&=\sum_{k=j-i}^\infty \frac{1}{k!}\binom{k}{j-i} \lambda_i^{k-(j-i)}T_1^k\\
%&=\frac{1}{(j-i)!}\sum_{k=j-i}^\infty \frac{1}{(k-(j-i))!} \lambda_i^{k-(j-i)}T_1^k\nonumber\\
%&=\frac{T_1^{j-i}e^{\lambda_i T_1}}{(j-i)!}\nonumber
%\end{align}
%which completes the proof for $gm(\lambda_i)>1$.
%\end{proof}

@@

@@

This induces two types of the errors detailed above: first, the truncation
error due to the domain transformations $\mathcal{F}_{\langle *
\rangle}(\cdot)$; and second, the rounding error introduced by the
multiplication operation.  We represent these errors as non-deterministic
additive noise.

@@

>>>>

In this work, equations~\eqref{eq:ode1} and~\eqref{eq:ode2} is soundly discretized in time (see Appendix~\ref{sec:cont_aa}) \todo{or use another paper as reference and remove sec .1?} into
\begin{align}
\label{eq:plant}
x_{k+1} = A_d x_k+ B_d u_k\\
y_{k} = C_d x_k + D_d u_k
\end{align} 
%
where $k \in \mathbb N$ and $x_{0}=x(0)$ is the initial state. 
$A_d$, $B_d$, $C_d$ and $D_d$ denote the matrices that describe the discretized plant dynamics, whereas $A$, $B$, $C$ and $D$ denote the continuous plant dynamics.  
%% We synthesize for requirements over this discrete-time domain. 
%% Later, we will address the issue of variable quantization, 
%% as introduced by the ADC/DAC conversion blocks (Fig.~\ref{fig:digitalsystem}).

 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 \subsubsection{Discrete dynamics using reals for systems without inputs}\label{sec:real_discrete_no_inputs}
 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 \todo{these theorems need more wording. what's their goal here?}
 \begin{theorem}
 Given $\dot{\vec{x}}=\mat{A}\vec{x}$, where $\mat{A}=\mat{S}\mat{J}\mat{S}^{-1}$, the expression
 \begin{align}
 \vec{x}_T&=\vec{x}(t=T)=\mat{A}_{T}\vec{x}_0\\
 \mat{A}_{T}&= \mat{S}
 \left [ \begin{array}{cccc}
 e^{T\lambda_1}  & s_1\frac{T^{1}e^{T\lambda_i}}{(1)!} & \hdots  & s_i\frac{T^{p-1}e^{T\lambda_i}}{(p-i)!} \\
0 & e^{T\lambda_i}  & s_i\frac{T^{j-i}e^{T\lambda_i}}{(j-i)!} & \vdots \\
\vdots & & \ddots & \vdots \\
0 & \cdots & 0  &e^{T\lambda_i} \\
\end{array} \right ]
 \mat{S}^{-1}
 \label{eq:continuous_tube_dyn}\\
 &\text{where } s_i=\left\{\begin{array}{cc}1&gm(\lambda_i)>1\\0&gm(\lambda_i)=1\end{array}\right.,\nonumber
 \end{align}
$\lambda_i \in \mat{J}$ are the eigenvalues of $\mat{A}$, and $gm(\lambda_i)$ is the geometric multiplicity of $\lambda_i$.  $\vec{x}_T$ is a witness of the system $\dot{\vec{x}}(t)=\mat{A}\vec{x}$ at time t=T.
 \end{theorem}
 \begin{corollary}
 If $\mat{J}$ is diagonal and there exists a discrete dynamics matrix for a sampling time $T_d$ such that $A_d=e^{\mat{A} T_d}$, then $\vec{x}_k=A_d^k\vec{x} : k \in (0,\infty)$ is a witness of the system at time $t=kT_d$.
 \end{corollary}
% \begin{proof}
% Let us recall equation \ref{eq:discretize}. The discrete representation of the system dynamics discretised with sample time $T_1$ is ruled by the formula
% $\mat{A}_1 = e^{\mat{A} T_1}$. From matrix theory, we have 
%\begin{align}
% e^{\mat{A}}&=\sum_{k=0}^\infty \frac{1}{k!}\mat{A}^k
%\end{align} 
%\begin{align} 
% e^{\mat{S}\mat{J}\mat{S}^{-1}}&=\sum_{k=0}^\infty \frac{1}{k!}\left(\mat{S}\mat{J}\mat{S}^{-1}\right)^k
% =\mat{S} \left (\sum_{k=0}^\infty \frac{1}{k!}\mat{J}^k\right) \mat{S}^{-1}\nonumber\\
% &=\mat{S}e^{\mat{J}}\mat{S}^{-1}
%\end{align} 
%\begin{equation}
%\mat{A}=\mat{S}\mat{J}\mat{S}^{-1} \Rightarrow \vec{A}_1 = \mat{S}\mat{J}_1\mat{S}^{-1}= \mat{S}e^{\mat{J} T_1}\mat{S}^{-1}.
% \end{equation}
% 
% Let $\mat{J}$ be a diagonal matrix, such that 
% $${\lambda_1}_i=e^{\lambda_i T_1}, \forall \lambda_i \in \mat{J}.$$
% Let us now take a different sample rate $T_2$, such that 
% $${\lambda_2}_i=e^{\lambda_i T_2}, \forall \lambda_i \in \mat{J}.$$
% We can then say that 
% \begin{equation}
% {\lambda_2}_i=e^{\lambda_i T_1 \frac{T_2}{T_1}}={\lambda_1}_i^{\frac{T_2}{T_1}} \Rightarrow A_2=A_1^{\frac{T_2}{T_1}}.
% \end{equation}
% This proves the theorem for $gm(\lambda_i)=1$.
% We now note that given an exponentiated Jordan form with geometric multiplicity, the upper diagonal terms can be
% referenced to the original eigenvalues modified by the sampling rate.
%  Let $\mat{J}\in \mathbb{R}^{p \times p}$ be a Jordan Canonical matrix.
% Then
%\begin{align}
% \mat{J}_1&=\sum_{k=0}^\infty \frac{1}{k!}\left(\mat{J}T_1\right)^k=\sum_{k=0}^\infty \frac{1}{k!} \left [ \begin{array}{cccc}
% \lambda_i^k  & \binom{k}{1}  \lambda^{k-1} & \hdots  & \binom{k}{p-1} \lambda_i^{k-p+1} \\
%& \lambda_i^k  & \binom{k}{1}  \lambda_i^{k-1} & \vdots \\
%\vdots & \ddots & \ddots & \vdots \\
%& &  &\lambda_i^k \\
%\end{array} \right ] T_1^k
%\end{align}
%The eigenvalues remain the same as in the diagonal case, but the upper triangular terms are of the form:
%\begin{align}
%\forall j>i, c_{ij}&=\sum_{k=j-i}^\infty \frac{1}{k!}\binom{k}{j-i} \lambda_i^{k-(j-i)}T_1^k\\
%&=\frac{1}{(j-i)!}\sum_{k=j-i}^\infty \frac{1}{(k-(j-i))!} \lambda_i^{k-(j-i)}T_1^k\nonumber\\
%&=\frac{T_1^{j-i}e^{\lambda_i T_1}}{(j-i)!}\nonumber
%\end{align}
%which completes the proof for $gm(\lambda_i)>1$.
%\end{proof}

@@

The models we consider have two sources of error that are due to numerical  
representation.  The first is the numerical error introduced by the
fixed-point numbers employed to model the plant, i.e., to represent the
plant dynamics $A_d$, $B_d$, $C_d$, $D_d$, $x_k$ and $y_k$.
The second is the quantization error
introduced by the digital controller, which performs operations on
fixed-point numbers.  In this section we outline the notation for the
representation of numbers, and briefly describe the errors
introduced.  A~formal discussion is in
Appendix~\ref{sec:numeric_rep}.

Let $\mathbb{R}_{\langle M,F \rangle}$ be a fixed point domain with $M$ bits
representing the binary word length of the representation and $F$ bits
representing the prescaling $2^{-F}$ (\emph{i.e} the number of bits after
the decimal place).  We note that a negative $F$ denotes an integer
multiplication that adds $-F$ trailing zeros in the number represented by
$M$.  Let also $\mathbb{R}_{\langle M,[E] \rangle}$ be a floating point
domain with $M$ bits mantissa and $E$ bits exponent.  This domain
corresponds to $$\mathbb{R}_{\langle M,[E] \rangle} = \bigcup_F
\mathbb{R}_{\langle M,F \rangle} \mid F \in [2^{1-E}, 2^{E-1}]$$ This
definition important because it means that the distance between any two
numbers in the domain depends on the location of those numbers.  Since
%we have shown that
a floating point domain is a union of a number of fixed
point domains, we will continue our reasoning on fixed point only, which
will apply to the floating point domain by extension.  Operations on these
numbers will incur errors that can be bound
by~\cite{DBLP:conf/arith/BrainTRW15}.

We will use the domain $\mathbb{R}_{\langle C \rangle} \mid \langle C \rangle = \langle M_c,F_c \rangle$
to represent the fixed point precision of the controller, and 
$\mathbb{R}_{\langle P \rangle} \supseteq \mathbb{R}_{\langle C \rangle}$
to represent the domain of the plant model.
Thus, any mathematical operations in our modelled digital controller will be executed in the
domain $\mathbb{R}_{\langle C \rangle}$, and all other calculations
in our model will be carried out in the domain $\mathbb{R}_{\langle
  P \rangle}$.
%% The physical plant operates in the reals, which means
%% our verification phase must also account for the numerical error caused by representing the
%% physical plant at the finite precision $\mathbb{R}_{\langle P \rangle}$.

%% We will use these subindices to represent functions, number, vectors and matrices in their
%% corresponding domains.

\todo{unclear}
We further note that by choosing $\langle C \rangle = \langle M_c,[E_c] \rangle$
we extend the following to floating point reasoning.

%% \subsubsection{Effect on safety specification and stability}

%% Let us first consider the effect of the quantization errors on safety. 
%% Within the controller, state values are manipulated at low precision,
%% alongside the vector multiplication $Kx$.
%% %
%% %computed considering the error introduced by the digital controller caused by performing the vector multiplication $Kx$ at low digital precision.
%% The inputs are computed using the following equation: 
%% %
%% \begin{align*}
%% u_{k}&=-(\mathcal{F}_{\langle I_c,F_c \rangle}(K)\cdot\mathcal{F}_{\langle I_c,F_c \rangle}(x_{k})). 
%% \end{align*}

%% This induces two types of the errors detailed above: first, the truncation
%% error due to representing $x_k$ as $\mathcal{F}{\langle I_c,F_c
%% \rangle}(x_{k})$; and second, the rounding error introduced by the
%% multiplication operation.  We represent these errors as non-deterministic
%% additive noise.%, which might de-stabilize the closed-loop model.

%% An additional error is due to the representation of the plant dynamics, namely 
%% %
%% \begin{align*}
%% x_{k+1} &=\mathcal{F}_{\langle I_p,F_p \rangle}(A_d) \mathcal{F}_{\langle I_p,F_p \rangle}(x_{k}) + \mathcal{F}_{\langle I_p,F_p \rangle}(B_d)\mathcal{F}_{\langle I_p,F_p \rangle}(u_{k}).
%% \end{align*}
%% We address this error by use of interval
%% arithmetic~\cite{moore1966interval} in the verification phase.

%% Previous studies~\cite{gangli1} show that the FWL affects the poles and
%% zeros positions, degrading the closed-loop dynamics, causing steady-state
%% errors (see Appendix~\ref{sec:appendix:LTIbackground} for details) and
%% eventually de-stabilizing the system~\cite{Bessa16}.  However, since in this
%% paper we require stability only as a precursor to safety, it is sufficient
%% to check that the (perturbed, noisy) model converges to a neighborhood of
%% the equilibrium within the safe set (see Appendix~\ref{sec:stab_FWL}).

%% In the following, we shall disregard these steady-state errors (caused by
%% FWL effects) when stability is ensured by synthesis, and then verify its
%% safety accounting for the finite-precision errors.

@@

This induces two types of the errors detailed above: first, the truncation
error due to the domain transformations $\mathcal{F}_{\langle *
\rangle}(\cdot)$; and second, the rounding error introduced by the
multiplication operation.  We represent these errors as non-deterministic
additive noise. More details can be found in Appendix~\ref{appendix:quantization}.

@@

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsubsection{Quantization error}
\label{appendix:quantization}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
